---
title: "Tidy eval Cheatsheet"
author: "Leonardo Uchôa Pedreira"
date: "1/13/2022"
output:
  pdf_document: default
urlcolor: blue
linkcolor: red
# header-includes:
#    - \usepackage{listings}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,warning = FALSE)
```

```{r libs,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
```

# General Recipe

Looking not deep, it's kinda simple to create new functions in the tidyverse way. The strategy to make it work is just to [`quote` (make it become a symbol) and `unquote` (evaluate the symbol)](https://tidyeval.tidyverse.org/sec-up-to-speed.html#quote-and-unquote) the function arguments. Let's look an example.

The next function call won't work.

```{r, eval=FALSE}
grouped_mean <- function(data, group_var, summary_var) {
    data %>%
    group_by(group_var) %>%
    summarise(mean = mean(summary_var))
}

grouped_mean(mtcars, cyl, mpg)

# Error: Must group by variables found in `.data`.
# * Column `group_var` is not found.
# Run `rlang::last_error()` to see where the error occurred.

```

To make it work we need to use the `quote` and `unquote` strategy. So first we quote the arguments with `enquo` and then unquote them with `!!`.

```{r, eval=TRUE}
grouped_mean <- function(data, group_var, summary_var) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)

  data %>%
    group_by(!!group_var) %>%
    summarise(mean = mean(!!summary_var))
}

grouped_mean(mtcars, cyl, mpg)
```

\pagebreak

### Some Extras

There some key functions when talking about tidy eval. They are:

- `quote`: used outside functions to make objects become symbols
- `enquo`: used inside functions to make objects become symbols. (There's also `enquos`).
- `sym`: used inside and outside functions to make strings become symbols. There's also `syms`, used for vectors, like `syms(c('cyl','disp'))`. Example:
  - this works `syms(c('cyl','disp'))`
  - this doesn't `sym(c('cyl','disp'))`
- `!!`: used to unquote symbols
- `!!!`: used to unquote vectorized symbols. For example:
  - this works `group_by(mtcars,!!!syms(c('cyl','disp')))`
  - this doesn't `group_by(mtcars,!!syms(c('cyl','disp')))`


### Using with `ggplot2`

If we're creating a plotting function like `plt <- function(df,col)`, we have two approaches. The first is to make `col` be a string and the second is to make it behave like an object, the dplyr style. 

\begin{center}
\textbf{First}: \textit{col} is called as a string
\end{center}

```{r, eval=FALSE}
col_summary <- function(df, col) {
  ggplot(df) + 
    geom_bar(aes(x = .data[[col]])) + 
    coord_flip()
}

col_summary(mpg, "drv")
```


\begin{center}
\textbf{Second}: \textit{col} is called like an object
\end{center}

```{r, eval=FALSE}
col_summary <- function(df, col) {
  ggplot(df) + 
    geom_bar(aes(x = {{ col }})) + 
    coord_flip()
}

col_summary(mpg, drv)
```

\pagebreak


# Sources

- [tidy eval book](https://tidyeval.tidyverse.org/sec-up-to-speed.html#writing-functions)
- [programming with dplyr](https://dplyr.tidyverse.org/articles/programming.html)
- [ggplot2 functions](https://ggplot2.tidyverse.org/articles/ggplot2-in-packages.html)
- [rlang cheatsheet](https://psych252.github.io/psych252book/figures/cheatsheets/tidyeval.pdf)
